Gaia identifies VFTS682 as the most massive runaway star known to date

### 2018-05-11 VFTS682.tex
incorporated Fabian's comments

\title{Gaia identifies VFTS682 as the most massive, dynamically ejected runaway star known to date}

\author{M.~Renzo$^{1}$, S.~E.~de~Mink$^{1}$, D.~J.~Lennon$^{2}$,
  S.~Justham$^{3,4}$, J.~Vink$^{5}$, A.~de~Koter$^{1}$ \todo{... add coauthors...} }

\affil{$^{1}$Astronomical Institute Anton Pannekoek, University of
    Amsterdam, 1098 XH Amsterdam, The Netherlands \\
    $^{2}$ ESA, European Space Astronomy Centre, Apdo. de Correos 78, E-28691 Villanueva de la Ca\~nada, Madrid, Spain \\
    $^{3}$ 
  % \and{School of Astronomy \& Space Science, University of the Chinese
  %   Academy of Sciences, Beijing 100012, China}
  % \and{National Astronomical Observatories, Chinese Academy of
  %   Sciences, Beijing 100012, China}
}

\thanks{M.~Renzo, \href{mailto:m.renzo@uva.nl}{m.renzo@uva.nl}}
\date{}
\begin{abstract}
The formation and fates of the most massive stars remains one of the most elusive questions in stellar astrophysics. 
%
Previous spectroscopic studies identified VFTS682 as one of the most massive stars known. It resides in the field of the 30 Dor region in the Large Magellanic Cloud at a projected distance of 29\,pc of the star cluster R136. The lack of a detected offset in radial velocity derived from the spectra, lead to speculations whether such extreme stars can form in relative isolation. 
%
We use Gaia's unprecedented proper motion measurements to investigate the kinematics of this object. We show that VFTS 682 is a bonafide runaway star moving with a relative velocity of about $60\kms$ away from the star cluster R136.  The inferred age, distance, and the magnitude and direction of the velocity are consistent a dynamical ejection scenario from the cluster. 
%
If correct, this would make VFTS 682 the most massive dynamically ejected runaway star known to date.  It proves that star clusters are capable of ejecting their most massive members. We discuss the implications for the existence of stars with masses well above the canonical upper mass limit inside the cluster, the internal dynamics in dense star clusters and the final fate of this object.

@@

How do massive stars form is one of the major longstanding questions in astrophysics
\citep[e.g.,][]{zinnecker:07}. Obtaining clues from observations has been challenging, because massive stars are intrinsically rare, 
%\citep[e.g.,][]{salpeter:55,kroupa:01, schneider:18}, 
evolve fast, typically reside in dense groups and they remain enshrouded in their parent cloud during the formation
process.  Major progress has been made on the theoretical side,
\citep[e.g.][]{kuiper:15,rosen:16}, but simulations cannot yet resolve
the smallest length scales that are of interest to understand the 
high level of multiplicity among massive stars  \citep[][]{sana:12,sana:17}.  Understanding massive star formation, and its possible dependence on environment and metallicity, is crucial for understanding the role massive stars play within their host galaxies as sources of feedback, but also for understanding the transients that mark their death and the compact remnants they leave behind.  This is therefore a key question for the present and upcoming transient surveys e.g., LSST and LIGO/Virgo,
%\citep[][]{}\todo{ref} . Blackgem is still hypothetical and a very tiny project compared to the others.  
which  will reveal transients associated to massive stars
evolution and death.

In has been proposed that most, if not all, stars form in clusters \citep{lada:03}, where massive stars are thought to  form preferentially in the dense cores of clusters. In this picture, field stars are primarily the result of the dissolution of dense groups. 
However, a significant population of massive stars exists in relative
isolation,  far from dense clusters or OB associations. The origin of
this population still remains debated \citep{lamb:16}.   One
hypothesis is that they  formed in the field, but this poses a
challenge for the theories of star formation.  The alternative
hypothesis is that these massive stars  formed in clusters, and then
been ejected from their birth locations.  Such ejections may result
from dynamical interactions \citep[e.g.,][]{poveda:67} or from the disruption of binary systems by the core-collapse of the companion star \citep[e.g.,][]{zwicky:57, blaauw:61}.

An interesting contribution to the debate on whether or not massive
stars an form in relative isolation was presented by
\cite{bestenlehner:11}, who discuss the case of the very massive star
VFTS682.   This star is located in the field of the 30 Doradus region
in the Large Magellanic Cloud (LMC) and was studied as part of the
multi-epoch spectroscopic VLT-FLAMES Tarantula Survey (VFTS) of
massive stars \citep{evans:11}.  Its spectral type is WNh5 and its 
present-day mass is $137.8^{+27.5}_{-15.9}\,M_\odot$,
\citep{schneider:18}. These correspond to an inferred initial mass of
$150.0^{+28.7}_{-17.4}\,M_\odot$ and a mass loss rate of
$\sim10^{-4.1\pm0.2}\,M_\odot \ \mathrm{yr}^{-1}$
\citep[][]{bestenlehner:11,schneider:18}, which make it one of the
most extreme objects in the region.  It is reminiscent of the very
massive stars, a.k.a. ``monster stars'', identified by
\citet{crowther:10, crowther:16} in the core of the
R136 cluster.

@@

Our results indicate that VFTS 682 is a bona fide runaway star.  The direction an magnitude of the velocity vector is fully consistent with dynamical ejection from R136.  Its inferred age is consistent with the inferred travel time. This means that the hypothesis that VFTS 682 is formed in relative isolation is rejected.  This makes VFTS 682, with an inferred present day mass well above a hundred solar masses, the most massive runaway star known to date.

@@

In this section we summarize the remarkable properties derived for VFTS682 \todo{?}\Secref{data:vfts683}
in earlier studies and we discuss we Gaia data in \Secref{data:gaia}.

@@

\citet{bestenlehner:11} further analyzed the spectra using the non-LTE
model atmosphere code CMFGEN \citep{hillier:98} to derive the stellar
parameters and surface abundances. They inferred a peculiar extinction
($R_V\simeq4.7$), leading to a surprisingly high luminosity
$\log_{10}(L/L_\odot) =  6.5\pm0.2$. \citep{schneider:18} estimated
the present-day mass of VFTS682 at  $137.8^{+27.5}_ {-15.9}\Msun$, an
apparent age of $1.0^{+0.2}_{-0.2}$\,Myr and an inferred initial mass
of $150.0^{+28.7}_{-17.4}\Msun$, using the Bayesian analysis tool Bonsai
\citep{schneider:17}, based on the evolutionary models from
\citet{brott:11, kohler:15}. These values place VFTS682 around the ``canonical upper limit'' of $\sim$$150\Msun$  by \citet{figer:05}.

@@

VFTS682 is identified with the id 4657685637907503744 in the Gaia DR2
catalog\footnote{\url{https://vizier.u-strasbg.fr/viz-bin/VizieR-3?-source=I/345/gaia2}}. Gaia
reports a G-band magnitude of 15.65. The star has a
\texttt{visibility\_period} = 17, which counts how many observations have
been used to reconstruct its astrometric solution
\citep[][]{lindengren:18}, and the reported
\texttt{astrometric\_excess\_noise} = 0. These values suggest that the Gaia
data for VFTS682 are trustworthy. However, the effective temperature
reported in Gaia DR2 is one order of magnitude lower than what found by
\cite{bestenlehner:11}, and the best fit parallax of this star is
negative. We do not use the effective temperature of the star anywhere
in this study, and we attribute the unphysical value of the parallax
to the large distance to the LMC. Our main findings do not rely on the
parallax nor the effective temperature values reported in the Gaia DR2
catalog.

@@

We calculate the averaged proper motion components for the whole
region using 
\begin{equation}
  \label{eq:mean}
  \langle \mu_i\rangle = \frac{\sum_\mathrm{stars}\frac{1}{\Delta
      \mu_i}\mu_i}{\sum_\mathrm{stars} \frac{1}{\Delta \mu_i}} \ \ , \
  \ \Delta \langle \mu_i\rangle = \frac{\sqrt{N}}{\sum_\mathrm{stars}
    \frac{1}{\Delta \mu_i}} \ \ ,
\end{equation}
where $i = \mathrm{RA}, \mathrm{DEC}$, and $\Delta \mu_i$ is the error
on the proper motion component reported by Gaia. The sums run over
all the $N=651$ stars in our selected sample. We evaluate each proper motion
component separately. For simplicity, throughout this study, we assume the same
distance of $50$\,kpc to the star \citep[][]{lebouteiller:08}, and to
the 30 Doradus region as a whole. We do not consider the error bars on
the distance determination when converting proper motions into
physical velocities. The data retrieved, and the ipython notebook used for the analysis
presented here will be made available at \todo{probably git repo on bitbucket?}.

@@

Using the 651 stars selected as described in \Secref{data:gaia} (a
subset is shown in blue in \Figref{fig:main}), we find averaged proper motion components of
$\langle\mu_\mathrm{RA}\rangle = 1.683\pm0.002\,\mathrm{mas\ yr^{-1}}$ and
$\langle\mu_\mathrm{DEC}\rangle =
0.672\pm0.003\,\mathrm{mas\ yr^{-1}}$. We note that these values are
in good agreement with what found by \cite{lennon:18}. Subtracting these values from the
proper motions of VFTS682 (see \Tabref{tab:vfts682}), we obtain the
components of proper motion of the star relative to the surrounding region
$\mu_\mathrm{RA} = 0.16\pm 0.07\,\mathrm{mas\ yr^{-1}}$ and $\mu_\mathrm{DEC} =
0.11\pm 0.08\,\mathrm{mas\ yr^{-1}}$. We note that the error budget is
dominated by the errors on the proper motion components of VFTS682.

These can be converted in the
the components of the relative transverse velocity $\delta v_\mathrm{RA}=38\pm17\,\kms$,
$\delta v_\mathrm{DEC}=26\pm19\,\kms$, assuming a distance of
50\,kpc (we do not account for the uncertainty in the distance
estimate when propagating errors). The radial velocity from
\cite{bestenlehner:11} then gives the third component along
the line of sight, allowing us to calculate the three-dimensional
peculiar speed of the star:

@@

The red arrow in \Figref{fig:main} shows the proper motion of VFTS682
relative to the region, and the lighter red arrows show the possible
range of directions within the uncertainties in the measured proper
motion. It is clear that the most likely origin of the star is R136,
as predicted from numerical N-body simulations by \cite{fujii:11, banerjee:12}.

@@

\begin{equation}
  \label{eq:kin_age}
  \tau_\mathrm{kin} = \frac{d_\parallel}{v_\parallel} \simeq
  \frac{29\,\mathrm{pc}}{46\,\kms} \simeq 0.63\, \mathrm{Myr} \ \ ,
\end{equation}
where $d_\parallel =29$\,pc is the projected distance from VFTS682 to
the core of R136 \citep[][]{bestenlehner:11}, $v_\parallel \equiv \sqrt{\left(\delta v_\mathrm{RA}\right)^2
    +\left(\delta v_\mathrm{DEC}\right)^2} =46\pm
17\,\kms$ is the speed of the star relative to the cluster projected on the sky, obtained
using the relative proper motion components calculated in
\Secref{sec:runaway}, and we use the approximation $1 \kms \simeq 1
\mathrm{pc \ Myr^{-1}}$.

@@

Based on our results, we can claim that VFTS682 is the most massive
runaway known to date, with a peculiar three-dimensional speed of $64\pm21\,\kms$. This means that isolated star formation is
\emph{not} required to explain this star. Its proper motion suggests that it was ejected from the cluster R136
$\sim$$0.6$\,Myr ago. Because of the exceptionally large mass
of this star, this raises the question of which stars must populate
the core of the cluster.

Dynamical ejections due to N-body interactions typically (although, not necessarily) eject the least
massive star among those interacting \todo{ref}. This means that, just
based on the kinematic properties of VFTS682, we would expect several
stars with initial masses larger than $\sim$$150\,M_\odot$ in the
cluster R136.
This can be considered an independent confirmation of the detection
of extremely massive stars by \cite{crowther:10} in the core of the
cluster. The projected rotational equatorial velocity of VFTS682
reported by \cite{schneider:18} is $v\sin(i)<200\,\kms$, which is in
line with the average rotation rate of massive stars in the region
\citep[][]{ramirez-agudelo:15}. This suggests that VFTS682 (i) has not
experienced rotationally induced chemically homogeneous evolution
\citep[][]{maeder:00,demink:09,marchant:16}, and (ii) it has not
accreted mass from a binary companion, nor it will, since the multi-epoch
data of the VFTS survey rule out the presence of a companion at
present day. Moreover, the spectral type of VFTS682
\citep[WNh5,][]{bestenlehner:11} is the same as R136a1-a3, i.e.~the
three most massive stars detected in the core of the cluster by
\cite{crowther:10}, with an astonishing similarity in particular with
the spectrum of R136a3. Therefore,
VFTS682 might be an ideal target to constrain the stellar physics of
stars with masses well above $\sim$$100\,M_\odot$: its isolation makes
it an easier target for observations compared to the similar stars
present in the crowded core of R136.

@@

\begin{acknowledgements}
  We are grateful to S.~Torres for 
  very useful discussions on the
  Gaia DR2 dataset, and to M.~C.~Ramirez-Tannus for invaluable help.
  SdM has received funding under the European Unions Horizon 2020 research and innovation programme from the European Research
  Council (ERC) (Grant agreement No. 715063).
  This work has made use of data from the ESA space mission Gaia (http://www.cosmos.esa.int/gaia), processed by the Gaia Data Processing and Analysis Consortium (DPAC, http://www.cosmos.esa.int/web/gaia/dpac/consortium). Funding for the DPAC has been provided by national institutions, in particular the institutions participating in the Gaia Multilateral Agreement. 
  The background image of \Figref{fig:main} is based on observations
  made with ESO Telescopes at the La Silla Observatory under programme
  ID 076.C-0888, processed and released by the ESO VOS/ADP group.
  This research made use of APLpy, an open-source plotting package for Python \citep[][]{robitaille:12}.
\end{acknowledgements}

>>>>

\title{Gaia identifies VFTS682 as the most massive % , dynamically
                                % ejected %% to fit in one line
  runaway star known to date}

\author{M.~Renzo$^{1}$, S.~E.~de~Mink$^{1}$, D.~J.~Lennon$^{2}$,
  S.~Justham$^{3,4}$, J.~Vink$^{5}$, F.~R.~Schneider$^6$, A.~de~Koter$^{1}$ \todo{... add coauthors...} }

\affil{$^{1}$Astronomical Institute Anton Pannekoek, University of
    Amsterdam, 1098 XH Amsterdam, The Netherlands \\
    $^{2}$ ESA, European Space Astronomy Centre, Apdo. de Correos 78, E-28691 Villanueva de la Ca\~nada, Madrid, Spain \\
    $^{3}$ {School of Astronomy \& Space Science, University of the Chinese
    Academy of Sciences, Beijing 100012, China}\\
     $^4$ {National Astronomical Observatories, Chinese Academy of
       Sciences, Beijing 100012, China}\\
     $^5$ {Armagh Observatory, College Hill, Armagh BT61 9DG, UK}\\
     $^6$ {Department of Physics, University of Oxford, Keble Road, Oxford OX1 3RH, UK}  
}

\thanks{M.~Renzo, \href{mailto:m.renzo@uva.nl}{m.renzo@uva.nl}}
\date{}
\begin{abstract}
The formation and fates of the most massive stars remains one of the most elusive questions in stellar astrophysics. 
%
Previous spectroscopic studies identified VFTS682 as one of the most massive stars known. It resides in the field of the 30 Dor region in the Large Magellanic Cloud at a projected distance of 29\,pc of the star cluster R136. The lack of a detected offset in radial velocity derived from the spectra, lead to speculations whether such extreme stars can form in relative isolation. 
%
We use the unprecedented proper motions from the second Gaia data
release (DR2) to investigate the kinematics of this object. We show
that VFTS 682 is a bonafide runaway star moving with a relative
velocity of about $64\pm21\kms$ away from the star cluster R136.  The
inferred age, distance, and the magnitude and direction of the
velocity are consistent with a dynamical ejection scenario from the cluster. 
%
This would make VFTS 682 the most massive runaway star known to date.  It proves that star clusters are capable of ejecting their most massive members. We discuss the implications for the existence of stars with masses well above $100\Msun$ inside the cluster, the internal dynamics in dense star clusters and the final fate of this object.

@@

How do massive stars form is one of the major longstanding questions in astrophysics
\citep[e.g.,][]{zinnecker:07}. Obtaining clues from observations has been challenging, because massive stars are intrinsically rare, 
%\citep[e.g.,][]{salpeter:55,kroupa:01, schneider:18}, 
evolve fast, typically reside in dense groups and they remain enshrouded in their parent cloud during the formation
process.  Major progress has been made on the theoretical side,
\citep[e.g.][]{kuiper:15,rosen:16}, but simulations cannot yet resolve
the smallest length scales that are of interest to understand the 
high level of multiplicity among massive stars  \citep[e.g.,][]{bate:09, sana:17}.  Understanding massive star formation, and its
possible dependence on environment and metallicity, is crucial for
understanding the role massive stars play within their host galaxies
as sources of feedback, but also for understanding the transients that
mark their death and the compact remnants they leave behind.  This is
therefore a key question for the present and upcoming transient
surveys e.g., Large Synoptic Spectroscopic Survey and LIGO/Virgo,
%\citep[][]{}\todo{ref} . Blackgem is still hypothetical and a very tiny project compared to the others.  
which  will reveal transients associated to massive stars
evolution and death.

In has been proposed that most, if not all, stars form in clusters \citep{lada:03}, where massive stars are thought to  form preferentially in the dense cores of clusters. In this picture, field stars are primarily the result of the dissolution of dense groups. 
However, a significant population of massive stars exists in relative
isolation,  far from dense clusters or OB associations. The origin of
this population still remains debated \citep{lamb:16,ward:18}.   One
hypothesis is that they  formed in the field, but this poses a
challenge for the theories of star formation. The alternative
hypothesis is that these massive stars formed in clusters, and then
been ejected from their birth locations. Such ejections may result
from dynamical interactions \citep[e.g.,][]{poveda:67} or from the disruption of binary systems by the core-collapse of the companion star \citep[e.g.,][]{zwicky:57, blaauw:61}.

An interesting contribution to the debate on whether or not massive
stars an form in relative isolation was presented by
\cite{bestenlehner:11, bressert:12}, who discuss the case of the very massive star
VFTS682.   This star is located in the field of the 30 Doradus region
in the Large Magellanic Cloud (LMC) and was studied as part of the
multi-epoch spectroscopic VLT-FLAMES Tarantula Survey (VFTS) of
massive stars \citep{evans:11}.  Its spectral type is WNh5 and its 
present-day mass is $137.8^{+27.5}_{-15.9}\,M_\odot$,
\citep{schneider:18}. These correspond to an inferred initial mass of
$150.0^{+28.7}_{-17.4}\,M_\odot$ and a wind mass loss rate of
$\sim10^{-4.1\pm0.2}\,M_\odot \ \mathrm{yr}^{-1}$ inferred from the
spectral analysis not accounting for clumping
\citep[][]{bestenlehner:11}, which make it one of the
most extreme objects in the region.  It is reminiscent of the very
massive stars % , a.k.a. ``monster stars'',
identified by
\citet{crowther:10, crowther:16} in the core of the
R136 cluster.

@@

Our results indicate that VFTS 682 is a bona fide runaway star.  The direction and magnitude of the velocity vector is fully consistent with dynamical ejection from R136.  Its inferred age is consistent with the inferred travel time. This means that the hypothesis that VFTS 682 is formed in relative isolation is rejected.  This makes VFTS 682, with an inferred present day mass well above a hundred solar masses, the most massive runaway star known to date.

@@

% feels like unnecessary here...
% In this section we summarize the remarkable properties derived for VFTS682 \todo{?}\Secref{data:vfts683}
% in earlier studies and we discuss we Gaia data in \Secref{data:gaia}.

@@

\citet{bestenlehner:11} further analyzed the spectra using the non-LTE
model atmosphere code CMFGEN \citep{hillier:98} to derive the stellar
parameters and surface abundances. They inferred a peculiar extinction
($R_V\simeq4.7$), leading to a surprisingly high luminosity
$\log_{10}(L/L_\odot) =  6.5\pm0.2$. \citep{schneider:18} estimated
the present-day mass of VFTS682 at  $137.8^{+27.5}_ {-15.9}\Msun$, an
apparent age of $1.0^{+0.2}_{-0.2}$\,Myr and an inferred initial mass
of $150.0^{+28.7}_{-17.4}\Msun$, using the Bayesian analysis tool Bonnsai
\citep{schneider:17}, based on the evolutionary models from
\citet{brott:11, kohler:15}. These values place VFTS682 around the
``canonical upper limit'' of $\sim$$150\Msun$ by \citet{figer:05}.

@@

VFTS682 is identified with the id 4657685637907503744 in the Gaia DR2
catalog\footnote{\url{https://vizier.u-strasbg.fr/viz-bin/VizieR-3?-source=I/345/gaia2}}. Gaia
reports a G-band magnitude of 15.65. The star has a
\texttt{visibility\_period} = 17, which counts how many observations have
been used to reconstruct its astrometric solution
\citep[][]{lindengren:18}, and the reported
\texttt{astrometric\_excess\_noise} = 0. These values suggest that the Gaia
data for VFTS682 are reliable. However, the effective temperature
reported in Gaia DR2 is one order of magnitude lower than what found by
\cite{bestenlehner:11}, and the best fit parallax of this star is
negative. We do not use the effective temperature of the star anywhere
in this study, and we attribute the unphysical value of the parallax
to the large distance to the LMC. Our main findings do not rely on the
parallax nor the effective temperature values reported in the Gaia DR2
catalog.

@@

We calculate the averaged proper motion components for the whole
region using 
\begin{equation}
  \label{eq:mean}
  \langle \mu_i\rangle = \frac{\sum_\mathrm{stars}\frac{1}{\Delta
      \mu_i}\mu_i}{\sum_\mathrm{stars} \frac{1}{\Delta \mu_i}} \ \ , \
  \ \Delta \langle \mu_i\rangle = \frac{\sqrt{N}}{\sum_\mathrm{stars}
    \frac{1}{\Delta \mu_i}} \ \ ,
\end{equation}
where $i = \mathrm{RA}, \mathrm{DEC}$, and $\Delta \mu_i$ is the error
on the proper motion component reported by Gaia. The sums run over
all the $N=651$ stars in our selected sample. We evaluate each proper motion
component separately. For simplicity, throughout this study, we assume the same
distance of $50$\,kpc to the star \citep[][]{pietrzynski:13}, and to
the 30 Doradus region as a whole. We do not consider the error bars on
the distance determination when converting proper motions into
physical velocities. The data retrieved, and the ipython notebook used for the analysis
presented here will be made available at \todo{probably git repo on bitbucket?}.

@@

Using the 651 stars selected as described in \Secref{data:gaia} (a
subset is shown in blue in \Figref{fig:main}), we find averaged proper motion components of
$\langle\mu_\mathrm{RA}\rangle = 1.683\pm0.002\,\mathrm{mas\ yr^{-1}}$ and
$\langle\mu_\mathrm{DEC}\rangle =
0.672\pm0.003\,\mathrm{mas\ yr^{-1}}$. We note that these values are
in good agreement with what is found by \cite{lennon:18}. Subtracting these values from the
proper motions of VFTS682 (see \Tabref{tab:vfts682}), we obtain the
components of proper motion of the star relative to the surrounding region
$\mu_\mathrm{RA} = 0.16\pm 0.07\,\mathrm{mas\ yr^{-1}}$ and $\mu_\mathrm{DEC} =
0.11\pm 0.08\,\mathrm{mas\ yr^{-1}}$. We note that the error budget is
dominated by the errors on the proper motion components of VFTS682.

These can be converted in the
the components of the relative transverse velocity $\delta v_\mathrm{RA}=38\pm17\,\kms$,
$\delta v_\mathrm{DEC}=26\pm19\,\kms$, assuming a distance of
50\,kpc. % (we do not account for the uncertainty in the distance
% estimate when propagating errors).
The radial velocity from
\cite{bestenlehner:11} then gives the third component along
the line of sight, allowing us to calculate the three-dimensional
peculiar speed of the star:

@@

The red arrow in \Figref{fig:main} shows the proper motion of VFTS682
relative to the region, and the lighter red arrows show the possible
range of directions within the uncertainties in the measured proper
motion. It is clear that the most likely origin of the star is R136,
as suggested based on numerical N-body simulations by \cite{fujii:11, banerjee:12}.

@@

\begin{equation}
  \label{eq:kin_age}
  \tau_\mathrm{kin} = \frac{d_\parallel}{v_\parallel} \simeq
  \frac{29\,\mathrm{pc}}{46\,\kms} \simeq 0.6\pm\,0.2 \mathrm{Myr} \ \ ,
\end{equation}
where $d_\parallel =29$\,pc is the projected distance from VFTS682 to
the core of R136 \citep[][]{bestenlehner:11}, $v_\parallel \equiv \sqrt{\left(\delta v_\mathrm{RA}\right)^2
    +\left(\delta v_\mathrm{DEC}\right)^2} =46\pm
17\,\kms$ is the speed of the star relative to the cluster projected on the sky, obtained
using the relative proper motion components calculated in
\Secref{sec:runaway}, and we use the approximation $1 \kms \simeq 1
\mathrm{pc \ Myr^{-1}}$. As in the rest of this study, we neglect for
simplicity the error on the distance estimates.

@@

Based on our results, we can claim that VFTS682 is the most massive
runaway known to date, with a peculiar three-dimensional speed of $64\pm21\,\kms$. This means that isolated star formation is
\emph{not} required to explain this star. Its proper motion suggests that it was ejected from the cluster R136
$0.6\pm0.2$\,Myr ago. Because of the exceptionally large mass
of this star, this raises the question of which stars must populate
the core of the cluster.

Dynamical ejections due to N-body interactions typically (although, not necessarily) eject the least
massive star among those interacting \todo{ref}. This means that, just
based on the kinematic properties of VFTS682, we would expect several
stars with initial masses larger than $\sim$$150\,M_\odot$ in the
cluster R136.
This is consistent with the detection
of extremely massive stars by \cite{crowther:10} in the core of the
cluster. The projected rotational equatorial velocity of VFTS682
reported by \cite{schneider:18} is $v\sin(i)<200\,\kms$, which is in
line with the average rotation rate of massive stars in the region
\citep[][]{ramirez-agudelo:15}. This suggests that VFTS682 (i) has not
experienced rotationally induced chemically homogeneous evolution
\citep[][]{maeder:00,demink:09,marchant:16}, and (ii) it has not
accreted mass from a binary companion, nor it will, since the multi-epoch
data of the VFTS survey rule out the presence of a companion at
present day. Moreover, the spectral type of VFTS682
\citep[WNh5,][]{bestenlehner:11} is the same as R136a1-a3, i.e.~the
three most massive stars detected in the core of the cluster by
\cite{crowther:10}, with an astonishing similarity in particular with
the spectrum of R136a3. Therefore,
VFTS682 might be an ideal target to constrain the stellar physics of
stars with masses well above $\sim$$100\,M_\odot$: its isolation makes
it an easier target for observations compared to the similar stars
present in the crowded core of R136.

@@

\begin{acknowledgements}
  \small
  We are grateful to S.~Torres for 
  very useful discussions on the
  Gaia DR2 dataset, and to M.~C.~Ramirez-Tannus for invaluable help.
  SdM has received funding under the European Unions Horizon 2020 research and innovation programme from the European Research
  Council (ERC) (Grant agreement No. 715063).
  This work has made use of data from the ESA space mission Gaia (\url{http://www.cosmos.esa.int/gaia}), processed by the Gaia Data Processing and Analysis Consortium (DPAC, \url{http://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding for the DPAC has been provided by national institutions, in particular the institutions participating in the Gaia Multilateral Agreement. 
  The background image of \Figref{fig:main} is based on observations
  made with ESO Telescopes at the La Silla Observatory under programme
  ID 076.C-0888, processed and released by the ESO VOS/ADP group.
  This research made use of APLpy, an open-source plotting package for Python \citep[][]{robitaille:12}.
\end{acknowledgements}
